Robust receding horizon control for linear systems with state and input dependent disturbances

### 2015-06-29 post_review.tex
Post-review edits

In all aforementioned work and, to the authors' best knowledge, in all of the published literature on algorithmic
solutions to the RMPC problem, the disturbances are confined to sets that are assumed to be constant,
i.e.~independent of the system state and control inputs. Recent algorithmic advances~\citep{Schaich:2015}
make it possible to treat polytopic disturbance sets that depend in an affine manner on the system state and the control
input. This provides an algorithmic and analytical framework which is used in the
current paper to embed affine state and input dependent disturbance constraints into the framework of the
active set RMPC solver framework proposed by~\cite{Buerger:2011}. Disturbances that belong to sets which depend
on model states and inputs can account for multiplicative model uncertainty as well as linearisation
errors, and the approach of this paper is therefore able to handle linearisation errors in RMPC non-conservatively.
\textcolor{red}{\cite{Rakovic:2006} proposes an algorithm with which reachable sets can be computed when the disturbance
is a realisation of a set defined on the product space of the disturbance the state and the input, i.e. rather
then an explicit dependence of the set on the state and input it is implicit. Hence the work discussed in
this paper is not covered by the analysis by~\cite{Rakovic:2006}.}

@@

\section{Problem Statement}\label{sec:problem:statement}
We consider a computationally efficient method for solving online min-max receding horizon control problems
that exploits the structure of multi-parametric programming solutions. Our approach determines the optimal
feedback law without approximating the cost or the constraints of the problem, and without making restrictive
assumptions on the parametrisation of control inputs.
%This broadens the applicability of the approach, allowing for higher order models and longer
%prediction horizons than 
%
%The high computational complexity of online min-max receding horizon control schemes generally requires either
%that the solution is approximated or that their application is limited to problems involving simple system models and short horizons. Here we consider an
%efficient method for solving min-max problems that exploits the structure of multi-parametric programming
%solutions.
%The set of systems to which a min-max robust receding horizon control scheme can be applied online 
%without approximating the solution is bounded by the system complexity. The way to solve min-max problems we 
%adopt here, is to use the structure of the solution to derive multi-parametric programs that can be solved
%efficiently. In order to be able to obtain a closed form solution of the min-max program simple optimisation
%formulations have to be
%The algorithmic treatment of min-max problems
%involving state constraints and disturbances constrained to lie in fixed polytopic disturbance sets was
%introduced in~\cite{Bertsekas:1971}. Recently~\cite{Schaich:2015} proposed an algorithm to compute the so called maximal
%robust positive invariant (MRPI) sets (see~\cite{blanchini:2007}) for linear systems subject to state and
%input dependent disturbance constraints. In this paper we develop the algorithm further
%to compute polytopic constraints necessary to satisfy state constraints robustly in this setup. We
%present a robust receding horizon control scheme for
%
To make it possible to obtain a closed form solution of the min-max optimal control problem we assume a
linear-quadratic problem formulation. Thus we assume linear dynamics 
\begin{equation}\label{eq:system:equation}
	x^+ = A x + B u + D w,
\end{equation}
%
where $A,B,D$ are real matrices of appropriate dimension, $x,  x^+\in\mathcal X\subseteq\mathbb R^n$ denote the state 
and successor state respectively, $u\in\mathcal U=\{u:F_iu\leq 1\,\forall i\in\mathcal I^{\mathcal U}\}
\subset\mathbb R^p$ is the control input
and $w\in\mathcal W(x,u)\subset\mathbb R^q$ is an unknown disturbance input, $\mathcal I^{\mathcal U}$ denotes
the index set of the constraint set~$\mathcal U$. The disturbance set $\mathcal
W(x,u)$ is assumed to have a known dependence on the values of $x$ and $u$, according to
%
\begin{equation}\label{eq:PWA:def:disturbance}
	\mathcal W(x,u) = \{w: Gw \leq H(x,u)\},
\end{equation}
%
where $H(x,u)$ is an elementwise convex, piecewise affine function. The control objective is to minimize the
worst case value, over all future model uncertainty, of the cost
\begin{equation}\label{eq:quadratic:cost}
\sum_{k=0}^\infty \bigl( \norm{x_k}_Q^2 + \norm{u_k}_R^2 - \gamma^2 \norm{w_k}^2 \bigr) ,
\end{equation}
where the subscript $k$ denotes the realisation of a variable at time $k$, $\norm{x}_Q^2$ denotes the quadratic
form $x^TQx$, $Q$ and $R$ are given positive definite
matrices and $\gamma$ is a given scalar weight that controls the $l^2$-gain of the closed loop system from the
disturbance sequence to the state and control sequences.

Given a prediction horizon of $N$ steps over which the predicted control law is to be optimized, we define the
min-max optimal control problem with $m$ stages to go as
%
\begin{subequations}\label{eq:the:problem}%
\begin{equation}
	J_{m}^\ast(x) = \min_u \max_{w,x^+} \tfrac{1}{2}\bigl(\norm{x}_Q^2+\norm{u}_R^2-\gamma^2
    \norm{w}^2\bigr) + J_{m-1}^\ast(x^+)
\end{equation}
subject to
\begin{gather}
	x^+ = A x + B u + D w\\
	u\in\mathcal U \\
	w\in\mathcal W(x,u) \\
	(x,u)\in\mathcal Z_m
\end{gather}
\end{subequations}
%
for $m=1,\ldots,N$, with $J_0^\ast(x) = \|x\|_{P_0}^2$.
The stagewise  state and input constraints $\mathcal Z_m$ are defined recursively by
%
\begin{subequations}\label{eq:definition:stage:constraints}
\begin{align}
	\mathcal Z_m &= \{(x,u)\,: u\in\mathcal U \nonumber\\
	&\quad\ \wedge \; A x + Bu + Dw\in\mathcal X_{m-1}\; \forall\, w\in\mathcal W(x,u)\} 
	\label{eq:definition:mixed:stage:constraint}\\
	\mathcal X_m &= \{x : \exists\, u\in\mathcal U, (x,u)\in\mathcal Z_m\}
	\label{eq:definition:state:stage:constraint}\\
\end{align}
with
\begin{equation}
	\mathcal X_0 = \mathcal X^\infty .
\end{equation}
\end{subequations}
Thus the constraints $(x,u)\in\mathcal Z_m$ ensure that the terminal state, namely $x^+$ at $m=1$, is
contained in $\mathcal{X}^\infty$. We define $\mathcal{X}^\infty$ as the maximal robust positive invariant 
(MRPI) set \citep{blanchini:2007}), for the closed-loop system governed by a fixed stabilising\NOTE{Really?! We don't really need it to be stabilising for this definition} state feedback 
controller $u=Kx$:
%
\begin{equation}\label{eq:definition:MRPI:set}
	\mathcal X^\infty = \{x: (A+BK)x + Dw \in \mathcal X^\infty \forall w\in\mathcal W(x,Kx)\},
\end{equation}
%
and we define $P_0$ so that $J_0^\ast(x)$ is the worst case value of the cost (\ref{eq:quadratic:cost}) under
$u=Kx$ in the absence of constraints.
Recently~\cite{Schaich:2015} proposed an algorithm to compute MRPI sets
for linear systems subject to state and input dependent disturbance constraints. In this paper we combine this approach
with the multistage multi-parametric programming method of~\cite{Buerger:ACC} to derive an efficient algorithmic
solution to the online MPC optimization of (\ref{eq:the:problem}a-e).

\section{Controllable sets}\label{sec:preliminaries}
In this section we summarise results on controllable sets that allow the constraints of the problem~\eqref{eq:the:problem} 
to be to formulated in a convenient way. The following result shows that the MRPI set (\ref{eq:definition:MRPI:set}) 
associated with the pointwise polytopic set~\eqref{eq:PWA:def:disturbance} is polytopic if $H$ is elementwise convex 
and piecewise affine. For a proof and detailed discussion of this result we refer the reader to~\cite{Schaich:2015}.
%
\begin{lem}\label{thm:MRPI:set:is:polytopic}
Let $\mathcal X \subseteq \{x:\Gamma x\leq{\bf{1}}\wedge-\Gamma x\leq{\bf{1}}\}$ be polyhedral, and the pair $(A+BK,\Gamma)$ be
observable, furthermore let $\mathcal W(x,Kx)$ be defined as in~\eqref{eq:PWA:def:disturbance} with a elementwise convex piecewise
affine function $H(x,Kx)$ which is bounded for all finite $x$. Then the MRPI set $\mathcal X^\infty$ is a polytope, i.e.
it has the representation 
\begin{equation}\label{eq:halfspace:representation:MRPI}
	\mathcal X^\infty=\{x:\Lambda_i x\leq \lambda_i, \forall\, i\in\mathcal I^\infty\}.
\end{equation}
\end{lem}
%
We need to emphasise the following 
\begin{rem}\label{rem:vertex:representation}
Note that an essential assumption in Lemma~\ref{thm:MRPI:set:is:polytopic} is that $\mathcal W(x,u)$ is a polytope
for all finite $(x,u)$, i.e. it has a convex hull representation $\mathcal W(x,u) = \conv_k\{w_k(x,u)\}, k\in\{1,\dots,r\}$ where each 
vertex, $w_k(x,u)$, is a piecewise affine function of $(x,u)$, i.e.
\begin{multline}
	w_k(x,u) = \max\{W_{k,1}^x x + W_{k,1}^u u + w_{k,1}, \\
        W_{k,2}^x x + W_{k,2}^u u + w_{k,2},\dots\}
\end{multline}
where the maximisation is performed elementwise.
\end{rem}
%
We now derive the representation of $\mathcal Z_m$ and $\mathcal X_m$ as defined in~\eqref{eq:definition:stage:constraints}.
According to Lemma~\ref{thm:MRPI:set:is:polytopic} the MRPI set $\mathcal X^\infty$ is polytopic and has the 
representation~\eqref{eq:halfspace:representation:MRPI}. Therefore the recursion~\eqref{eq:definition:mixed:stage:constraint}
defines $\mathcal Z_1$ as
\begin{equation}\label{eq:derivation:of:mpLP:stage:constraints}
\begin{aligned}
	\mathcal Z_1 =& (\mathbb R^n\times\mathcal U)\cap 
\begin{aligned}[t] \Bigl\{(x,u): \Lambda_{0,i}(Ax&+Bu+Dw)\leq \lambda_{0,i} \\
& \forall w\in\mathcal W(x,u),\, i\in\mathcal I_0 \Bigr\} \end{aligned}
\\
	=&\begin{aligned}[t] (\mathbb R^n & \times \mathcal U) \cap\Bigl\{ (x,u): \\
& \max_{w\in\mathcal W(x,u)}\Lambda_{0,i}(Ax+Bu+Dw)\leq\lambda_{0,i},\, \forall i\in\mathcal I_0 \Bigr\} \end{aligned}
\\
	=&\begin{aligned}[t] & (\mathbb R^n \times\mathcal U)\cap \Bigl\{ (x,u): \\
& \!\!\!\!\!\!
\Lambda_{0,i}(Ax+Bu)+ \!\!\underbrace{\max_{j\in\{1,\dots,r\}}\!\Lambda_{0,i} Dw_j(x,u)}_{(i)}\leq\lambda_i, \, \forall i\in\mathcal I_0
	\Bigr\} \end{aligned}
\end{aligned}
\end{equation}
Since the maximisation $(i)$ in \eqref{eq:derivation:of:mpLP:stage:constraints} is a multi-parametric
linear program, its solution is piecewise affine in $(x,u)$. 
Therefore the set $\mathcal Z_1$ has a representation
$\mathcal Z_1=\{(x,u):\Xi^x_{1,i}x+\Xi^u_{1,i}u\leq\xi_{1,i},i\in\mathcal J_1\}$, with~$\mathcal J_1$ denoting
its index set. The projection~$\mathcal X_1$ 
of $\mathcal Z_1$ onto~$\mathbb R^n$ therefore has the representation~$\mathcal X_1 = \{x:\Lambda_{1,i} x\leq\lambda_{1,i},i\in\mathcal I_1\}$.
From here the determination of $\mathcal Z_k$ and $\mathcal X_k$ follows recursively. We are now able to formulate 
all constraints in~\eqref{eq:the:problem} as inequalities. In the next section we discuss a method of solving
min-max programs such as~\eqref{eq:the:problem} efficiently.

@@

	(f^T+\theta^TS^T)N_\beta \neq 0

@@

and can be made arbitrarily small by the choice of $\beta$. Therefore we restrict $\theta$ to satisfy 

@@

any purpose. For a general mpQP, condition~\eqref{mpQP:compatability:assumption} is a restriction on the set
of parameters for which~\eqref{mpQP:primal} has a solution of the form~\eqref{mpQP:solution}.
In our multi-stage min-max setup, the parameter is the optimiser of a subsequent optimisation
problem and can be ensured to satisfy~\eqref{mpQP:compatability:assumption} by the so called 
compatibility constraint. The optimal objective of~\eqref{mpQP:primal} is given by

@@

\section{Exact solution of min-max program}\label{sec:line:search}
The previous section presents a recursive algorithm to solve~\eqref{eq:the:problem}
when all active constraints are known. In this section we present a method to find the active
constraints for a given initial state $\chi_e$ using a line search starting from a state $\chi_0$
for which the active constraints are known. In Sections~\ref{sec:gen:mpQP} and~\ref{sec:applied:mpQP}
we derived an explicit solution for each optimisation sub-problem which depends affinely on the
state~$x$ at that stage~$m$ or the state and the input at that stage. Henceforth we denote the stage 
with an index, e.g.~$x_m$ denotes the optimised state in stage $m$, that is $x^{+,\ast}$ at stage $m+1$.
To ensure continuity of the dual variables we choose $\hat\zeta_m = \hat\beta_m$ and $\zeta_m=\beta_{m-1}$.
The affine character of the solution of each sub-problem allows us to parametrise the entire trajectory
with respect to parameters at stage~$m=N$, that is $x_N=\chi_0$ and $\beta_N$.

We recursively define all primal and dual variables:
\begin{equation}\label{eq:trajectory:affine:representation}
	\begin{split}
		u_N(x_N) =& \underbrace{K_{x,N}^u}_{\Psi_N^u} x_N + \underbrace{k_N^u}_{\psi_N^u}\\
		\nu_N(x_N,\beta_N) &= \underbrace{L_{x,N}^\nu}_{\Psi_N^\nu} x_N + \underbrace{L_{\beta,N}^\nu}_{\Phi_N^\nu} \beta_N + 
		\underbrace{l_N^\nu}_{\psi_N^u}\\
		&\vdots \\
		x_0(x_N) =& \underbrace{M_{z,1}^{x^+}\left(\begin{array}{c} \Psi_2^x \\ \Psi_1^u \end{array}\right)}_{\Psi_1^x}x_N + 
		\underbrace{M_{z,1}^{x^+}\left(\begin{array}{c} \psi_2^x \\ \psi_1^u \end{array}\right) + m_1^{x^+}}_{\psi^x_1}\\
		\eta_1(x_N,\beta_N) =& \underbrace{\left(N_{z,1}^{\eta}\left(\begin{array}{c} \Psi_2^x \\ \Psi_1^u \end{array}\right) + 
		N_{\hat\beta,1}^\eta \Psi_1^{\hat\zeta}\right)}_{\Psi_1^\eta}x_N + \underbrace{N_{\hat\beta,1}^\eta \Phi_1^{\hat\zeta}}_{\Phi_1^\eta} \beta_N\\
		&+\underbrace{N_{z,1}^{\eta}\left(\begin{array}{c} \psi_2^x \\ \psi_1^u \end{array}\right) + 
		N_{\hat\beta,1}^\eta \psi_1^{\hat\zeta} + n_1^\eta}_{\psi_1^\eta}\\
		&\vdots
	\end{split}
\end{equation}
Notice that $\Phi^{(\cdot)}_m$ is only non-trivial if the associated sub-problem is degenerate. The
representation~\eqref{eq:trajectory:affine:representation} is valid as long as the constraint set remains
unchanged, i.e.\ the inequalities that hold with equality do not change.
\textcolor{red}{With~\eqref{eq:trajectory:affine:representation} we can justify the proposed choice of the degeneracy parameters
$\beta_m$ and $\hat\beta_m$. When constraints are activated their dual variable starts at zero and changes
in an affine way in the appropriate direction, at the moment a linearly dependent constraint is activated, i.e. the
constraints become rank deficient, three things happen: The dual variable to the degenerate constraint suddenly
depends on an additional degeneracy parameter, a compatibility constraint is introduced in the predecessor
problem and that compatibility constraint's dual variable is given as a affine function starting at zero,
the remaining problem is not affected by the degeneracy. By the proposed choice we couple
the degeneracy variable to dual variable of the associated compatibility constraint, hence adding an 
additional affine function to the already affine representation of the dual variable and thereby ensuring continuity
of the dual variable.}

>>>>

\edit{In almost all existing methods for RMPC, the model disturbances are assumed to be confined to sets that are constant, i.e.~independent of the system state and control inputs. Disturbances that depend on model states (or control inputs) are considered in~\cite{Kuntsevich:1995,Rakovic:2006}. However general formulations of state and input dependence can result in non-convex constraints~\citep{Rakovic:2006}. To avoid this we make use of recent algorithmic advances~\citep{Schaich:2015} that enable a class of polytopic disturbance sets that depend affinely on the system state and control input to be handled using convex constraints. This provides an algorithmic and analytical framework which is used in the
current paper to embed affine state and input dependent disturbance constraints into the framework of the
active set RMPC solver framework proposed by~\cite{Buerger:2011}. Disturbances that belong to sets which depend
on model states and inputs can account for multiplicative model uncertainty as well as linearisation
errors, and the approach of this paper is therefore able to handle linearisation errors in RMPC non-conservatively.}

% \edit{In all aforementioned work and, to the authors' best knowledge, in all of the published literature on algorithmic
% solutions to the RMPC problem, the disturbances are confined to sets that are assumed to be constant,
% i.e.~independent of the system state and control inputs. Recent algorithmic advances~\citep{Schaich:2015}
% make it possible to treat polytopic disturbance sets that depend in an affine manner on the system state and the control
% input. This provides an algorithmic and analytical framework which is used in the
% current paper to embed affine state and input dependent disturbance constraints into the framework of the
% active set RMPC solver framework proposed by~\cite{Buerger:2011}. Disturbances that belong to sets which depend
% on model states and inputs can account for multiplicative model uncertainty as well as linearisation
% errors, and the approach of this paper is therefore able to handle linearisation errors in RMPC non-conservatively.}
% \textcolor{red}{\cite{Rakovic:2006} proposes an algorithm with which reachable sets can be computed when the disturbance
% is a realisation of a set defined on the product space of the disturbance the state and the input, i.e. rather
% then an explicit dependence of the set on the state and input it is implicit. Hence the work discussed in
% this paper is not covered by the analysis by~\cite{Rakovic:2006}.}

@@

\section{Problem Statement}\label{sec:problem:statement}
We consider a computationally efficient method for solving online min-max receding horizon control problems
that exploits the structure of multi-parametric programming solutions. Our approach determines the optimal
feedback law without approximating the cost or the constraints of the problem, and without making restrictive
assumptions on the parametrisation of control inputs.
%This broadens the applicability of the approach, allowing for higher order models and longer
%prediction horizons than 
%
%The high computational complexity of online min-max receding horizon control schemes generally requires either
%that the solution is approximated or that their application is limited to problems involving simple system models and short horizons. Here we consider an
%efficient method for solving min-max problems that exploits the structure of multi-parametric programming
%solutions.
%The set of systems to which a min-max robust receding horizon control scheme can be applied online 
%without approximating the solution is bounded by the system complexity. The way to solve min-max problems we 
%adopt here, is to use the structure of the solution to derive multi-parametric programs that can be solved
%efficiently. In order to be able to obtain a closed form solution of the min-max program simple optimisation
%formulations have to be
%The algorithmic treatment of min-max problems
%involving state constraints and disturbances constrained to lie in fixed polytopic disturbance sets was
%introduced in~\cite{Bertsekas:1971}. Recently~\cite{Schaich:2015} proposed an algorithm to compute the so called maximal
%robust positive invariant (MRPI) sets (see~\cite{blanchini:2007}) for linear systems subject to state and
%input dependent disturbance constraints. In this paper we develop the algorithm further
%to compute polytopic constraints necessary to satisfy state constraints robustly in this setup. We
%present a robust receding horizon control scheme for
%
To make it possible to obtain a closed form solution of the min-max optimal control problem we assume a
linear-quadratic problem formulation. Thus we assume linear dynamics 
\begin{equation}\label{eq:system:equation}
	x^+ = A x + B u + D w,
\end{equation}
%
where $A,B,D$ are real matrices of appropriate dimension, $x,  x^+\in\mathcal X\subseteq\mathbb R^n$ denote the state 
and successor state respectively, $u\in\mathcal U=\{u:F_iu\leq 1\,\forall i\in\mathcal I^{\mathcal U}\}
\subset\mathbb R^p$ is the control input
and $w\in\mathcal W(x,u)\subset\mathbb R^q$ is an unknown disturbance input, $\mathcal I^{\mathcal U}$ denotes
the index set of the constraint set~$\mathcal U$. The disturbance set $\mathcal
W(x,u)$ is assumed to have a known dependence on the values of $x$ and $u$, according to
%
\begin{equation}\label{eq:PWA:def:disturbance}
	\mathcal W(x,u) = \{w: Gw \leq H(x,u)\},
\end{equation}
%
\edit{where each element of $H(x,u)$ is a convex, piecewise affine function for all $(x,u)\in\mathcal X\times\mathcal U$}. 
%\edit{where $H(x,u)$ is an elementwise convex, piecewise affine function}.
The control objective is to minimize the
worst case value, over all future model uncertainty, of the cost
\begin{equation}\label{eq:quadratic:cost}
\sum_{k=0}^\infty \bigl( \norm{x_k}_Q^2 + \norm{u_k}_R^2 - \gamma^2 \norm{w_k}^2 \bigr) ,
\end{equation}
where the subscript $k$ denotes the realisation of a variable at time $k$, $\norm{x}_Q^2$ denotes the quadratic
form $x^TQx$, $Q$ and $R$ are given positive definite
matrices and $\gamma$ is a given scalar weight that controls the $l^2$-gain of the closed loop system from the
disturbance sequence to the state and control sequences.

Given a prediction horizon of $N$ steps over which the predicted control law is to be optimized, we define the
min-max optimal control problem with $m$ stages to go as
%
\begin{subequations}\label{eq:the:problem}%
\begin{equation}
	J_{m}^\ast(x) = \min_u \max_{w,x^+} \tfrac{1}{2}\bigl(\norm{x}_Q^2+\norm{u}_R^2-\gamma^2
    \norm{w}^2\bigr) + J_{m-1}^\ast(x^+)
\end{equation}
subject to
\begin{gather}
	x^+ = A x + B u + D w\\
	u\in\mathcal U \\
	w\in\mathcal W(x,u) \\
	(x,u)\in\mathcal Z_m
\end{gather}
\end{subequations}
%
for $m=1,\ldots,N$, with $J_0^\ast(x) = \|x\|_{P_0}^2$.
The stagewise  state and input constraints $\mathcal Z_m$ are defined recursively by
%
\begin{subequations}\label{eq:definition:stage:constraints}
\begin{align}
	\mathcal Z_m &= \{(x,u)\,: u\in\mathcal U \nonumber\\
	&\quad\ \wedge \; A x + Bu + Dw\in\mathcal X_{m-1}\; \forall\, w\in\mathcal W(x,u)\} 
	\label{eq:definition:mixed:stage:constraint}\\
	\mathcal X_m &= \{x : \exists\, u\in\mathcal U, (x,u)\in\mathcal Z_m\}
	\label{eq:definition:state:stage:constraint}\\
\end{align}
with
\begin{equation}
	\mathcal X_0 = \mathcal X^\infty .
\end{equation}
\end{subequations}
Thus the constraints $(x,u)\in\mathcal Z_m$ ensure that the terminal state, namely $x^+$ at $m=1$, is
contained in $\mathcal{X}^\infty$. We define $\mathcal{X}^\infty$ as the maximal robust positive invariant 
(MRPI) set \citep{blanchini:2007}), for the closed-loop system
governed by a given
%stabilising\NOTE{Really?! We don't really need it to be stabilising for this definition} 
state feedback controller $u=Kx$:
%
\begin{equation}\label{eq:definition:MRPI:set}
	\mathcal X^\infty = \{x: (A+BK)x + Dw \in \mathcal X^\infty \ \forall w\in\mathcal W(x,Kx)\},
\end{equation}
%
and we define $P_0$ so that $J_0^\ast(x)$ is the worst case value of the cost (\ref{eq:quadratic:cost}) under
$u=Kx$ in the absence of constraints.
Recently~\cite{Schaich:2015} proposed an algorithm to compute MRPI sets
for linear systems subject to state and input dependent disturbance constraints. In this paper we combine this approach
with the multistage multi-parametric programming method of~\cite{Buerger:ACC} to derive an efficient algorithmic
solution to the online MPC optimization of (\ref{eq:the:problem}a-e).

\section{Controllable sets}\label{sec:preliminaries}
In this section we summarise results on controllable sets that allow the constraints of the problem~\eqref{eq:the:problem} 
to be formulated in a convenient way. The following result shows that the MRPI set (\ref{eq:definition:MRPI:set}) 
associated with the pointwise polytopic set~\eqref{eq:PWA:def:disturbance} is polytopic if $H$ is elementwise convex 
and piecewise affine. For a proof and detailed discussion of this result we refer the reader to~\cite{Schaich:2015}.
%
\edit{\begin{thm}[\cite{Schaich:2015}]\label{thm:MRPI:set:is:polytopic}
Let $\mathcal X \subseteq \{x:\Gamma x\leq{\bf{1}}\wedge-\Gamma x\leq{\bf{1}}\}$ be a polyhedral set, let the pair $(A+BK,\Gamma)$ be
observable, and let $\mathcal W(x,Kx)$ be defined as in~\eqref{eq:PWA:def:disturbance} with an elementwise convex piecewise
affine function $H(x,Kx)$ which is bounded for all finite $x$. Then the MRPI set $\mathcal X^\infty$ is polytopic, i.e.\ it has the representation 
\begin{equation}\label{eq:halfspace:representation:MRPI}
	\mathcal X^\infty=\{x:\Lambda_i x\leq \lambda_i, \forall\, i\in\mathcal I^\infty\}.
\end{equation}
\end{thm}
%
%We need to emphasise the following 
\begin{rem}\label{rem:vertex:representation}
The conditions of Theorem~\ref{thm:MRPI:set:is:polytopic} require that $\mathcal W(x,u)$ is polytopic
for all finite $(x,u)$, i.e.\ it must have a convex hull representation $\mathcal W(x,u) = \conv_k\{w_k(x,u)\}, k\in\{1,\dots,r\}$ where each 
vertex, $w_k(x,u)$, is a piecewise affine function of $(x,u)$, i.e.
\begin{multline}
	w_k(x,u) = \max\{W_{k,1}^x x + W_{k,1}^u u + w_{k,1}, \\
        W_{k,2}^x x + W_{k,2}^u u + w_{k,2},\dots\}
\end{multline}
where the maximisation is performed elementwise.
\end{rem}}
%
We now derive the representation of $\mathcal Z_m$ and $\mathcal X_m$ as defined in~\eqref{eq:definition:stage:constraints}.
According to Theorem~\ref{thm:MRPI:set:is:polytopic} the MRPI set $\mathcal X^\infty$ is polytopic and has the 
representation~\eqref{eq:halfspace:representation:MRPI}. Therefore the recursion~\eqref{eq:definition:mixed:stage:constraint}
defines $\mathcal Z_1$ as
\begin{equation}\label{eq:derivation:of:mpLP:stage:constraints}
\begin{aligned}
	\mathcal Z_1 =& (\mathbb R^n\times\mathcal U)\cap 
\begin{aligned}[t] \Bigl\{(x,u): \Lambda_{0,i}(Ax&+Bu+Dw)\leq \lambda_{0,i} \\
& \forall w\in\mathcal W(x,u),\, i\in\mathcal I_0 \Bigr\} \end{aligned}
\\
	=&\begin{aligned}[t] (\mathbb R^n & \times \mathcal U) \cap\Bigl\{ (x,u): \\
& \max_{w\in\mathcal W(x,u)}\Lambda_{0,i}(Ax+Bu+Dw)\leq\lambda_{0,i},\, \forall i\in\mathcal I_0 \Bigr\} \end{aligned}
\\
	=&\begin{aligned}[t] & (\mathbb R^n \times\mathcal U)\cap \Bigl\{ (x,u): \\
& \!\!\!\!\!\!
\Lambda_{0,i}(Ax+Bu)+ \!\!\underbrace{\max_{j\in\{1,\dots,r\}}\!\Lambda_{0,i} Dw_j(x,u)}_{(i)}\leq\lambda_i, \, \forall i\in\mathcal I_0
	\Bigr\} \end{aligned}
\end{aligned}
\end{equation}
Since the maximisation $(i)$ in \eqref{eq:derivation:of:mpLP:stage:constraints} is a multi-parametric
linear program, its solution is piecewise affine in $(x,u)$. 
Therefore the set $\mathcal Z_1$ has a representation
$\mathcal Z_1=\{(x,u):\Xi^x_{1,i}x+\Xi^u_{1,i}u\leq\xi_{1,i},i\in\mathcal J_1\}$, with~$\mathcal J_1$ denoting
its index set. The projection~$\mathcal X_1$ 
of $\mathcal Z_1$ onto~$\mathbb R^n$ therefore has the representation~$\mathcal X_1 = \{x:\Lambda_{1,i} x\leq\lambda_{1,i},i\in\mathcal I_1\}$.
From here the determination of $\mathcal Z_k$ and $\mathcal X_k$ follows recursively. We are now able to formulate 
all constraints in~\eqref{eq:the:problem} as inequalities. In the next section we discuss a method of solving
min-max programs such as~\eqref{eq:the:problem} efficiently.

@@

	(f^T+\theta^TS^T)N_\beta \neq 0 .

@@

%and can be made arbitrarily small by the choice of $\beta$. 
Therefore we restrict $\theta$ to satisfy 

@@

any purpose. \edit{For a general mpQP, the condition~\eqref{mpQP:compatability:assumption} ensures that~\eqref{mpQP:primal}
% is a restriction on the set
% of parameters $\theta$ for which~\eqref{mpQP:primal} 
has a solution of the form~\eqref{mpQP:solution}.
In our multi-stage min-max setup, the parameter $\theta$ is the optimiser of a subsequent optimisation
problem; hence~\eqref{mpQP:compatability:assumption} can be enforced by imposing it as a constraint (which we call a compatibility constraint) on $\theta$ in that problem.}
%the so called compatibility constraint. 
The optimal objective of~\eqref{mpQP:primal} is given by

@@

\section{Exact solution of min-max program}\label{sec:line:search}
The previous section presented a recursive algorithm to solve~\eqref{eq:the:problem}
when all active constraints are known. In this section we present a method to find the active
constraints for a given initial state $\chi_e$ using a line search starting from a state $\chi_0$
for which the active constraints are known. In Sections~\ref{sec:gen:mpQP} and~\ref{sec:applied:mpQP}
we derived an explicit solution for each optimisation sub-problem which depends affinely on the
state~$x$ at that stage~$m$ or the state and the input at that stage. Henceforth we denote the stage 
with an index, e.g.~$x_m$ denotes the optimised state in stage $m$, that is $x^{+,\ast}$ at stage $m+1$.
To ensure continuity of the dual variables we choose $\hat\zeta_m = \hat\beta_m$ and $\zeta_m=\beta_{m-1}$.
The affine character of the solution of each sub-problem allows us to parametrise the entire trajectory
with respect to parameters at stage~$m=N$, that is $x_N=\chi_0$ and $\beta_N$.

We recursively define all primal and dual variables:
\begin{equation}\label{eq:trajectory:affine:representation}
	\begin{split}
		u_N(x_N) =& \underbrace{K_{x,N}^u}_{\Psi_N^u} x_N + \underbrace{k_N^u}_{\psi_N^u}\\
		\nu_N(x_N,\beta_N) &= \underbrace{L_{x,N}^\nu}_{\Psi_N^\nu} x_N + \underbrace{L_{\beta,N}^\nu}_{\Phi_N^\nu} \beta_N + 
		\underbrace{l_N^\nu}_{\psi_N^u}\\
		&\vdots \\
		x_0(x_N) =& \underbrace{M_{z,1}^{x^+}\left(\begin{array}{c} \Psi_2^x \\ \Psi_1^u \end{array}\right)}_{\Psi_1^x}x_N + 
		\underbrace{M_{z,1}^{x^+}\left(\begin{array}{c} \psi_2^x \\ \psi_1^u \end{array}\right) + m_1^{x^+}}_{\psi^x_1}\\
		\eta_1(x_N,\beta_N) =& \underbrace{\left(N_{z,1}^{\eta}\left(\begin{array}{c} \Psi_2^x \\ \Psi_1^u \end{array}\right) + 
		N_{\hat\beta,1}^\eta \Psi_1^{\hat\zeta}\right)}_{\Psi_1^\eta}x_N + \underbrace{N_{\hat\beta,1}^\eta \Phi_1^{\hat\zeta}}_{\Phi_1^\eta} \beta_N\\
		&+\underbrace{N_{z,1}^{\eta}\left(\begin{array}{c} \psi_2^x \\ \psi_1^u \end{array}\right) + 
		N_{\hat\beta,1}^\eta \psi_1^{\hat\zeta} + n_1^\eta}_{\psi_1^\eta}
%\\
%		&\vdots
	\end{split}
\end{equation}
Notice that $\Phi^{(\cdot)}_m$ is only non-trivial if the associated sub-problem is degenerate. The
representation~\eqref{eq:trajectory:affine:representation} is valid as long as the constraint set remains
unchanged, i.e.\ the inequalities that hold with equality do not change.
\textcolor{red}{With~\eqref{eq:trajectory:affine:representation} we can justify \edit{choosing the degeneracy parameters as
$\beta_m=\zeta_{m+1}$ and $\hat\beta_m=\hat\zeta_m$. When a constraint is activated the associated dual variable starts at zero and depends affinely on $(x_N,\beta_N)$.
%changesin an affine way in the appropriate direction, 
When the active constraints of a sub-problem become linearly dependent three things happen: the dual variables to the degenerate constraints
depend on an additional degeneracy parameter, a compatibility constraint is introduced in the predecessor
problem, and that compatibility constraint's dual variable is given as a affine function starting at zero,
while all other sub-problems are not affected by the degeneracy. Through} the proposed choice we couple
the degeneracy variable to dual variable of the associated compatibility constraint, hence adding an 
additional affine function to the already affine representation of the dual variable and thereby ensuring continuity
of the dual variable.}

Given the known set of active
constraints at $\chi_0$ we perform a line search towards $\chi_e$ via  
$x_N(\alpha)=\chi_0+ \alpha(\chi_e-x_0)$ for $0\leq\alpha\leq1$. We
use~\eqref{eq:trajectory:affine:representation} to check all inactive constraints with the primal variables
$x_m(x_N(\alpha)),\, u_m(x_N(\alpha))$ and~$w_m(x_N(\alpha))$, and all dual variables corresponding to
inequality constraints to check whether constraints need to be deactivated. That is, we search for the
smallest $\alpha$ such that $\Xi_{m,i}^x x_m(x_N(\alpha))+\Xi_{m,i}^u u_m(x_N(\alpha))=\xi_{m,i}$ or
$G_iw_m(x_N(\alpha))=1$ or $\eta_{m,i}(x_N(\alpha),\beta_N)=0$ for some $m$ and $i$, and update the set of active
constraints accordingly, i.e. if an additional constraint is activated it is introduced to the equality
constraint mpQP recursion and the updated problems are solved again recursively, if a constraint is
deactivated because the smallest $\alpha$ corresponds to the zero-crossing of its dual variable then it is
removed and the
mpQP recursion is solved again. The updated active set is the new known active set at $\chi_0 = x_N(\alpha)$
and the line-search continues until the state~$\chi_e$ is in the same active set as $\chi_N(\alpha)$ where the
line-search terminates.  To initialise we start at the origin $\chi_0=0$, since we know that no constraints
are active.
